Further Evidence of Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution

### 2022-03-21 src/ms.tex
detection bias discussion

\edit1{\citet{vanSaders2019} addressed this issue in detail, finding that a threshold of Ro$_\mathrm{thresh} \approx 2$ could reproduce the observed upper edge of the \citet{McQuillan2014} \teff–\prot distribution. However, those authors presented two arguments that suggest detection bias is not solely responsible for producing the observed \teff–\prot distribution. First, stars at the upper edge are not those with the lowest-amplitude signals detected, as one would expect from a detection edge. For the variability amplitudes typical of Sun-like stars at the upper edge, one finds cooler stars with similar amplitude at even longer periods, as well as hotter stars with much smaller amplitudes (though at shorter periods). Even among Sun-like stars, those authors found that 9\% of stars below the edge have amplitudes smaller than stars with similar \teff at the edge (possibly due to issues such as flux dilution).}

\edit1{Critically, we also note that while an Ro-dependent detection threshold may reproduce the upper edge of the \teff–\prot distribution, it would not be responsible for producing a pile-up or overdensity at this edge. To our knowledge, there are no systematics of the rotation period catalogs considered here that would bias stars towards the detection threshold. We also note that the asteroseismic method of measuring rotation periods has separate observational biases, yet still seemingly yields a pile-up of stars at the long-period edge (as discussed in \S\ref{sec:asteroseismic}).}

>>>>

\edit1{\citet{vanSaders2019} addressed this issue in detail, finding that a threshold of Ro$_\mathrm{thresh} \approx 2$ could reproduce the observed upper edge of the \citet{McQuillan2014} \teff–\prot distribution. In this scenario, the detection threshold sets the upper edge of the observed \teff–\prot distribution at high \teff, while the age of the galaxy sets the upper edge at low \teff (since low-mass stars have not yet had enough time to spin down to Ro$_\mathrm{thresh}$). However, those authors presented arguments that suggest detection bias is not solely responsible for producing the observed \teff–\prot distribution. First, stars at the upper edge are not those with the lowest-amplitude signals detected, as one would expect from a detection edge. For the variability amplitudes typical of Sun-like stars at the upper edge, one finds cooler stars with similar amplitude at even longer periods, as well as hotter stars with much smaller amplitudes (though at shorter periods). Even among Sun-like stars, those authors found that 9\% of stars below the edge have amplitudes smaller than stars with similar \teff at the edge (possibly due to issues such as flux dilution).}

\edit1{Critically, we also note that while an Ro-dependent detection threshold may reproduce the upper edge of the \teff–\prot distribution, it would not be responsible for producing a pile-up or overdensity at this edge. To our knowledge, there are no systematics of the rotation period catalogs considered here that would bias stars towards the detection threshold. Furthermore, two independent samples with different observational biases than the rotation period catalogs considered here also appear to yield a pile-up of stars at the long-period edge. These samples are the Kepler asteroseismic sample, which measures rotation through asteroseismic mode splitting (discussed in \S\ref{subsec:asteroseismic}), and the sample of Kepler stars with \vsini from high-resolution spectroscopy, from which \citet{Masuda2021} inferred the peak of the \teff–\prot distribution using constraints on the stellar radii.}
